---
  title: "Fish Microplasstics"
author: "Ashlee Pilon + Maison"
date: "`r Sys.Date()`"
output:
  html_document: 
  theme: spacelab
---
  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library(ggmap)
library(lubridate)
library(shiny)
library(dplyr)
library(shinydashboard)
library(ggplot2)
library(readr)
``` 

download the data
```{r}
microplastics <- fish_microplastics
```

Clean names
```{r}
microplastics <- microplastics %>%
  clean_names()
```

Check NAs
```{r}
microplastics  %>%
  gg_miss_var()
```
Make a map of the locations where fish were collected from.
```{r}
register_stadiamaps("280bdc76-265c-4420-af16-f3fc7ab98211", write = FALSE)
```

```{r}
microplastics %>%
  select(collection_lat, collection_long) %>%
  summary()
```

```{r}
lat <- c(27, 27.75)
long <- c(-83, -82)
bbox <- make_bbox(long, lat, f= 0.03)
unique(microplastics$collection_long)
```

```{r}
map_fish <- get_stadiamap(bbox, maptype = "stamen_terrain", zoom = 7)
```

```{r}
ggmap(map_fish) +
  geom_point(data = microplastics, aes(collection_long, collection_lat), size = 3, color = "blue" , alpha = 0.8)+ 
  labs(x = "Longitude", y = "Latitude", title = "Collection Locations")
```
Make each location that fish were collected from (long values) into a new row 
```{r}
microplastics <- microplastics %>%
  mutate(station = case_when(
    collection_long == -82.64988 ~ "Station 1",
    collection_long == -82.54724 ~ "Station 2"
  ))
```

Mutate the fish_id column to create a new column called species that has the full name of the species of fish collected. The first two letters of the fish_id correspond to the species of fish collected.
```{r}
microplastics <- microplastics %>%
  mutate(code = substr(fish_id, 1, 2),
         species = case_when(
           code == "Oc" ~ "orthopristis chrysoptera",
           code == "OC" ~ "orthopristis chrysoptera",
           code == "Lr" ~ "lagodon rhomboides",
           code == "Af" ~ "ariopsis felis",
           code == "Ob" ~ "opsanus beta"
         ))
```

make the blank box in particle_type into NA
```{r}
microplastics <- microplastics %>%
  mutate(particle_type = na_if(particle_type, ""))
```

***Save our edited data frame as a new csv fil
```{r}
#write_csv(microplastics, "microplastics.csv"))

```
   
Build shiny app
```{r, eval=F, echo=T}
ui <- dashboardPage(
  
  skin = "purple",
  
  dashboardHeader(title = "Fish Microplastics"),
  
  dashboardSidebar(
    
    selectInput("z",
                "Select a variable of interest",
                choices = c("color", "particle_type", "sample_type", "species"),
                selected = "color"),
    
    selectInput("x",
                "Select a variable of interest",
                choices = c("station", "species"),
                selected = "station")
  ),
  
  dashboardBody(
    
    plotOutput("plot", width = "500px", height = "400px")
    
  )
)

server <- function(input, output, session) {

  output$plot <- renderPlot({
    
    microplastics %>%
      ggplot(aes(x = .data[[input$x]], fill = .data[[input$z]])) +
      geom_bar(position = "dodge") +
      theme_bw() +
      labs(x = input$x, fill = input$z)
    
  })
}

shinyApp(ui, server)
```

bar plot for count by station
```{r}
microplastics %>%
  ggplot(aes(x= station, fill = color)) +
  geom_bar(position = "dodge") +
  theme_bw() +
  labs(x = "Station", y = "Count of Fish Collected", title = "Count of Fish Collected by Station")
```


Boxplot of sample mass by species
```{r}
microplastics %>%
  ggplot(aes(x= species, y = net_sample_mass_g)) +
  geom_boxplot(fill = "aquamarine", color = "black") +
  theme_bw() +
  labs(x = "Species", y = "Sample Mass (g)", title = "Boxplot of Sample Mass by Species")
```
Boxplot of sample mass by station
```{r}
microplastics %>%
  ggplot(aes(x= station, y = net_sample_mass_g)) +
  geom_boxplot(fill = "aquamarine", color = "black") +
  theme_classic() +
  labs(x = "Species", y = "Sample Mass (g)", title = "Boxplot of Sample Mass by Species")
```
box plot of weights of each particle type 
```{r}
microplastics %>%
  filter(!is.na(particle_type)) %>%
  ggplot(aes(x= particle_type, y = net_sample_mass_g)) +
  geom_boxplot(fill = "aquamarine", color = "black") +
  theme_classic() +
  labs(x = "Particle Type", y = "Sample Mass (g)", title = "Boxplot of Sample Mass by Particle Type")
```

Particles taken from each sample type by species
```{r}
microplastics %>%
  ggplot(aes(x = sample_type, fill= sample_type)) + 
  geom_bar() + 
  theme_bw()+ 
  facet_wrap(~ species)+ 
  labs(title = "Muscle vs GI Sampling By Species", x = NULL, y = "Particle Count") 
```

Particle Types by Species
```{r}
microplastics %>%
  group_by(species, particle_type) %>%
  summarise(total = sum(quantity)) %>%
  filter(!is.na(particle_type)) %>%
  ggplot(aes(x = particle_type, y = total, fill = particle_type)) +
  geom_col() +
  facet_wrap(~ species) +
  labs(title = "Most Common Particle Type by Species", x = NULL, y = "Count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Everything was confirmed plastic
```{r}
microplastics %>%
  filter(!is.na(hot_needle), !is.na(particle_type)) %>%
  group_by(particle_type, hot_needle) %>%
  summarise(total = sum(quantity)) %>%
  ggplot(aes(x = reorder(particle_type, -total), y = total, fill = hot_needle)) +
  geom_col(position = "stack") +
  labs(title = "Confirmed vs Unconfirmed Plastic by Particle Type", x = NULL, y = "Count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Where plastic was found for each species in each sample
```{r}
microplastics %>%
  group_by(species, sample_type) %>%
  summarise(total = sum(quantity)) %>%
  filter(!is.na(sample_type)) %>%
  ggplot(aes(x = species, y = total, fill = species)) +
  geom_col() +
  facet_wrap(~ sample_type) +
  coord_flip() +         
  labs(title = "Plastic Consumed by Species and Tissue", x = NULL, y = "Count") +
  theme_bw()
```

Percentage of fihs with TWP by species
```{r}
microplastics %>%
  group_by(fish_id, species) %>%
  summarise(has_twp = any(particle_type == "Fragment_TWP")) %>%
  group_by(species) %>%
  summarise(pct = mean(has_twp) * 100) %>%
  ggplot(aes(x = species, y = pct, fill = species)) +
  geom_col() +
  labs(title = "% of Fish with Tire Wear Particles by Species", x = NULL, y = "%") +
  theme_bw()
```

